---
title: "p8105_hw6_map2365"
author: "Megan Panier"
date: "`r Sys.Date()`"
output: html_document
---

# Problem 2 
```{r data, include=FALSE}
#creating data directory
dir.create("data", showWarnings = FALSE)

url <- "https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv"
download.file(url, destfile = "data/homicide-data.csv", mode = "wb")

# Loading libraries
library(tidyverse)
library(janitor)

# cleaning data
homicide_data <- read.csv("data/homicide-data.csv") %>%
    janitor::clean_names() %>%
    mutate(
        city_state = paste(city, state, sep = ", "),
        status = ifelse(disposition %in% c("Closed without arrest", "Open/No arrest"), 0, 1)
    ) %>%
    filter(
        !city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL"),
        victim_race %in% c("White", "Black"),
        !is.na(victim_age)
    ) %>%
    mutate(victim_age = as.numeric(victim_age))
```
The dataset has `r nrow(homicide_data)` observations and `r ncol(homicide_data)` variables. The variables include `r colnames(homicide_data)`.

```{r}
#logistic regression for baltimore
baltimore_data <- homicide_data %>%
  filter(city_state == "Baltimore, MD")

baltimore_model <- glm(
  status ~ victim_age + victim_sex + victim_race,
  data = baltimore_data,
  family = binomial(link = "logit")
)

baltimore_results <- broom::tidy(baltimore_model, conf.int = TRUE, exponentiate = TRUE) %>%
  filter(term == "victim_sexMale") %>%
  select(estimate, conf.low, conf.high)

#logistic regression - all cities 
city_results <- homicide_data %>%
  group_by(city_state) %>%
  nest() %>%
  mutate(
    model = map(data, ~ glm(
      status ~ victim_age + victim_sex + victim_race,
      data = .x,
      family = binomial(link = "logit")
    )),
    tidy_model = map(model, ~ broom::tidy(.x, conf.int = TRUE, exponentiate = TRUE))
  ) %>%
  unnest(cols = tidy_model) %>%
  filter(term == "victim_sexMale") %>%
  arrange(estimate) %>%
  mutate(city_state = fct_reorder(city_state, estimate))

```
The adjusted odds ratio for the city of Baltimore is `r baltimore_results$estimate` with a confidence interval of (`r baltimore_results$conf.low`, `r baltimore_results$conf.high`). The odds of having a solved homicide case for those who were male is 0.42 times the odds for a female having a solved homicde case. We are 95% confident that the true estimate is somewhere between (`r baltimore_results$conf.low`, baltimore_results$conf.high`).

```{r}
#data visualization 

ggplot(city_results, aes(x = estimate, y = city_state)) +
  geom_point(color = "blue", size = 3) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2, color = "gray") +
  labs(
    title = "Adjusted Odds Ratios (OR) for Solving Homicides by City",
    x = "Odds Ratio (Male vs Female Victims)",
    y = "City"
  ) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "red") +
  theme_minimal(base_size = 12) +
  theme(axis.text.y = element_text(size = 8))

```
As seen in the visual above, in a vast majority of cities, there are higher odds that a homicide involving women will be solved. The exceptions to this are Stockton, CA, Fresno, CA, and Albuquerque, NM.






