---
title: "p8105_hw6_map2365"
author: "Megan Panier"
date: "`r Sys.Date()`"
output: html_document
---

# Problem 2 
```{r data, include=FALSE}
#creating data directory
dir.create("data", showWarnings = FALSE)

url <- "https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv"
download.file(url, destfile = "data/homicide-data.csv", mode = "wb")

# Loading libraries
library(tidyverse)
library(janitor)

# cleaning data
homicide_data <- read.csv("data/homicide-data.csv") %>%
    janitor::clean_names() %>%
    mutate(
        city_state = paste(city, state, sep = ", "),
        status = ifelse(disposition %in% c("Closed without arrest", "Open/No arrest"), 0, 1)
    ) %>%
    filter(
        !city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL"),
        victim_race %in% c("White", "Black"),
        !is.na(victim_age)
    ) %>%
    mutate(victim_age = as.numeric(victim_age))
```
The dataset has `r nrow(homicide_data)` observations and `r ncol(homicide_data)` variables. The variables include `r colnames(homicide_data)`.

```{r, warning=FALSE}
#logistic regression for baltimore
baltimore_data <- homicide_data %>%
  filter(city_state == "Baltimore, MD")

baltimore_model <- glm(
  status ~ victim_age + victim_sex + victim_race,
  data = baltimore_data,
  family = binomial(link = "logit")
)

baltimore_results <- broom::tidy(baltimore_model, conf.int = TRUE, exponentiate = TRUE) %>%
  filter(term == "victim_sexMale") %>%
  select(estimate, conf.low, conf.high)

#logistic regression - all cities 
city_results <- homicide_data %>%
  group_by(city_state) %>%
  nest() %>%
  mutate(
    model = map(data, ~ glm(
      status ~ victim_age + victim_sex + victim_race,
      data = .x,
      family = binomial(link = "logit")
    )),
    tidy_model = map(model, ~ broom::tidy(.x, conf.int = TRUE, exponentiate = TRUE))
  ) %>%
  unnest(cols = tidy_model) %>%
  filter(term == "victim_sexMale") %>%
  arrange(estimate) %>%
  mutate(city_state = fct_reorder(city_state, estimate))

```
The adjusted odds ratio for the city of Baltimore is `r baltimore_results$estimate` with a confidence interval of (`r baltimore_results$conf.low`, `r baltimore_results$conf.high`). The odds of having a solved homicide case for those who were male is 0.42 times the odds for a female having a solved homicde case. We are 95% confident that the true estimate is somewhere between (`r baltimore_results$conf.low`, baltimore_results$conf.high`).

```{r}
#data visualization 

ggplot(city_results, aes(x = estimate, y = city_state)) +
  geom_point(color = "blue", size = 3) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2, color = "gray") +
  labs(
    title = "Adjusted Odds Ratios (OR) for Solving Homicides by City",
    x = "Odds Ratio (Male vs Female Victims)",
    y = "City"
  ) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "red") +
  theme_minimal(base_size = 12) +
  theme(axis.text.y = element_text(size = 8))

```
As seen in the visual above, in a vast majority of cities, there are higher odds that a homicide involving women will be solved (indicated by the OR values being less than 1). The exceptions to this are Stockton, CA, Fresno, CA, and Albuquerque, NM (indicated by the OR>1). There are a couple of cities where the OR is ~1.0 which indicates that the odds for homicides involving males are approximately as likely to be solved as homicides involving females.  


# Problem 3 

```{r}
#data import and cleaning
birthweight <- read.csv("data/birthweight.csv", 
                        header = TRUE, 
                        sep = ",", 
                        na = c("NA", "")) |>
  janitor::clean_names() |>
  mutate(
    babysex = factor(babysex, levels = c(1, 2), labels = c("Male", "Female")),
    frace = factor(frace, levels = c(1, 2, 3, 4, 8, 9), labels = c("White", "Black", "Asian", "Puerto Rican", "Other", "Unknown")),
    mrace = factor(mrace, levels = c(1, 2, 3, 4, 8), labels = c("White", "Black", "Asian", "Puerto Rican", "Other")),
    malform = factor(malform, levels = c(0, 1), labels = c("No", "Yes"))
  ) |>
  drop_na()

# used to check data and missing values 
summary(birthweight)
colSums(is.na(birthweight))

```

```{r}
formula1 <- bwt ~ momage + ppbmi + gaweeks + babysex
formula2 <- bwt ~ blength + gaweeks
formula3 <- bwt ~ bhead * blength * babysex
```

formula1 = incorporates a subset of the available variables that are most likely to influence birth weight. For this first model, I chose to focus on key maternal factors (BMI before pregnancy and maternal age) as well as key factors for the fetus (including their gestational age and sex). I chose to only include these factors because I did not want an overly complicated model and I focused on variables that are know to affect birthweigth (based on prior literature). 

forumla2 = simpler model using length at birth and gestational age as predictors (main effects only)

formula3 = using head circumference, length, sex, and all interactions (including the three-way interaction) between these

```{r}
#M1

model1 <- lm(formula1, data = birthweight)
model1_pred <- birthweight %>%
  modelr::add_predictions(model1) %>%
  modelr::add_residuals(model1)

ggplot(model1_pred, aes(x = pred, y = resid)) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(
    x = "Fitted Values",
    y = "Residuals",
    title = "Model 1 Residuals vs. Fitted Values (M1)"
  ) +
  theme_minimal()
```
From the plot, we see that the data does not appear to be evenly distributed as there is some clustering around 0. Therefore, this model might might not fully capture the variability of birthweight for some of the predicted ranges. 
```{r}
model2 <- lm(formula2, data = birthweight)
model2_pred <- birthweight %>%
  modelr::add_predictions(model2) %>%
  modelr::add_residuals(model2)

ggplot(model2_pred, aes(x = pred, y = resid)) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(
    x = "Fitted Values",
    y = "Residuals",
    title = "Model 2 Residuals vs. Fitted Values (M2)"
  ) +
  theme_minimal()
```

```{r}
model3 <- lm(formula3, data = birthweight)
model3_pred <- birthweight %>%
  modelr::add_predictions(model3) %>%
  modelr::add_residuals(model3)

ggplot(model3_pred, aes(x = pred, y = resid)) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(
    x = "Fitted Values",
    y = "Residuals",
    title = "Model 3 Residuals vs. Fitted Values (M3)"
  ) +
  theme_minimal()
```
For each of the models, residuals were plotted against fitted values using modelr::add_predictions() and modelr::add_residuals(). I used this to evaluate the fit of the 3 models. Based on the plots, I see that all three models show approximately random scatter of points around the zero line and no obvious patterns/trends in the residuals. Furthermore, based on the plots for the 3 models, model 3 appears to have the best model fit. Model 3 appears to be the best model because it has smaller residuals and a more consistent variance pattern. Overall, from the plots, I would day that model 3 > model 2 > model 1 in terms of overall model fit. 





```{r}
# Monte Carlo cross validation
cv_df <- crossv_mc(birthweight, n = 100, test = 0.2) %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  ) %>%
  mutate(
    fit_model1 = map(train, ~ lm(formula1, data = .x)),
    fit_model2 = map(train, ~ lm(formula2, data = .x)),
    fit_model3 = map(train, ~ lm(formula3, data = .x))
  ) %>%
  mutate(
    rmse_model1 = map2_dbl(fit_model1, test, ~ rmse(model = .x, data = .y)),
    rmse_model2 = map2_dbl(fit_model2, test, ~ rmse(model = .x, data = .y)),
    rmse_model3 = map2_dbl(fit_model3, test, ~ rmse(model = .x, data = .y))
  )

# RMSE plot
cv_df %>%
  select(starts_with("rmse")) %>%
  pivot_longer(
    everything(),
    names_to = "model",
    values_to = "rmse",
    names_prefix = "rmse_"
  ) %>%
  mutate(model = fct_inorder(model)) %>%
  ggplot(aes(x = model, y = rmse)) +
  geom_violin() +
  labs(
    x = "Model",
    y = "RMSE",
    title = "Cross-Validation RMSE for the 3 Models"
  ) +
  theme_minimal()

```









